---
title: "Predictive Maching Learning Project"
author: "Dan Goodman"
date: "Friday, June 20, 2014"
---

This is my attempt at the PML project.  I was only able to generate working models with the "Accelerometer" predictors.

**Steps in Analysis**:

- Load training and testing datasets
- Filter out messy data-factors with only 2 levels
- Select only predictors starting with "Accel"
- Break training data into 2 partitions
- Create a CART model using caret
- Evaluate Performance against the testing partition
- Run model against test data for submission

Load Libraries
```{r echo=FALSE}

library(caret)
library(rpart.plot)
library(rattle)
library(AppliedPredictiveModeling)
library(ggplot2)
setwd("~/R/Coursera")

#Get data and clean up column types
training <- read.csv("pml-training.csv")
training$classe <- factor(training$classe)
testing <- read.csv("pml-testing.csv")
testing$var_accel_forearm <- as.numeric(testing$var_accel_forearm)
testing$var_total_accel_belt <- as.numeric(testing$var_total_accel_belt)
testing$var_accel_arm  <- as.numeric(testing$var_accel_arm)
testing$var_accel_dumbbell <- as.numeric(testing$var_accel_dumbbell)
ColFilter <- as.list(apply(training, 2, function(x)length(unique(x))))
Keep <- names(which(ColFilter>2))
training <- subset(training, select = Keep)
ColFilterTest <- as.list(apply(training, 2, function(x)length(unique(x))))
KeepTest <- Keep
KeepTest[153]="problem_id"
testing <- subset(testing, select = KeepTest)
#Split Training Set
inTrain = createDataPartition(training$classe, p = 0.7)[[1]]
TrainSet = training[ inTrain,]
TestSet = training[-inTrain,]

#Data is Messy, Focus on Accelerometers only
#Just Use Accelerometer Values
Accel <- grep("^accel", names(TrainSet))
AccelTrain <- TrainSet[,c(Accel,which(names(TrainSet)=="classe"))]
AccelTest <- TestSet[,c(Accel,which(names(TestSet)=="classe"))]


```
Variables in my dataset:
```{r}
names(AccelTest)

```

Create a Simple Tree
```{r}
#Try a Tree Model
ModelCart <- train(classe~., method="rpart", data=AccelTrain)
print(ModelCart)


```
Visualize tree
```{r plot}

fancyRpartPlot(ModelCart$finalModel)
```

Of the 13 predictors in the set, 4 are shown as useful.

-----
Compare against Test Set
```{r}

#Predict on TestSet
AccelPredict <- predict(ModelCart, newdata = AccelTest)
print(confusionMatrix(AccelPredict,AccelTest$classe))

```

The model generates an out of sample error rate of about 60%
It does reasonably well at Class A and Class E predictions, but needs some help on B, C, and D.

I'm sure this could be improved by expanding the list of predictors.